---
title: "APM466 A1"
author: "Keying Chen"
date: "2/5/2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 2.1 Fundamental Questions
#### 1.

(a)
Governments issue bonds to raise money from investors on its own credit in order to operate new projects or infrastructure, however, simply printing more money will speed up inflation which will cause the money in our hand become worthless.

(b) 
A yield curve might flatten in the long-term part which indicates that the investors and traders concerns about the macroeconomic outlook such as recession.

(c) 
Quantitative easing is that central banks trends to increase money supply by buying long-term government bonds or other securities, one of the examples is that the Fred has employed the quantitative easing by purchasing the long-term government bonds and other securities open-ended since the beginning of the COVID-19 pandemic.


#### 2

Since I need to construct a "0-5 year" yield & spot curves, I choose 10 bonds based on the issue date and maturity date. These 10 bonds that I choose are issued after 2018 and the maturity date of them are evenly distributed between 2022 to 2027 (5 years).

CAN 1.25 March 25, CAN 0.50 Sept 25, CAN 0.25 Mar 26, CAN 1.00 Sept 26, CAN 1.25 Mar 27, CAN 0.25 Aug 22, CAN 0.25 Feb 23, CAN 2.25 Mar 24, CAN 0.25 Aug 23, CAN 1.50 Oct 24

#### 3

The stochastic processes give us several random points that are evenly distributed along the curve, the covariance matrix of those stochastic processes represent the covariance between each two features. The eigenvectors and eigenvalues associated with the covariance matrix tell us the directions of new axis and the their relative magnitudes in this new feature space. The eigenvectors which have the lowest eigenvalues will be dropped since they do not make a big difference to the overall data. Principle Component Analysis aims to decrease the dimensionality of dataset without lossing much information.

## 2.2 Empirical Questions

#### 4.

#### (a) 


```{r, echo=FALSE}
library(readxl)
APM_A1_DATA <- read_excel("APM A1 DATA.xlsx")
ytm.matrix = as.matrix(APM_A1_DATA)
year = c(0.5,1,1.5,2,2.5,3,3.5,4,4.5,5)
plot(year,ytm.matrix[,2],type="o", main="5-year Yield Curve", col="black", xlab="Year", ylab="YTM(%)",ylim=c(0.6,1.8))
colour = c("red","orange","yellow","green","blue","purple","pink","navy","light blue")
for (i in c(3:11)){
   lines(year,ytm.matrix[,i], type="o", col=colour[i-1])
}
legend("bottomleft",legend=c("01/10/2022","01/11/2022","01/12/2022","01/13/2022","01/14/2022","01/17/2022","01/18/2022","01/19/2022","01/20/2022","01/21/2022"), col=c("black","red","orange","yellow","green","blue","purple","pink","navy","light blue"), lty=1, cex=0.8)
```



To calculate the bonds' yield(YTM), I use the equation $$P=\sum_{i=1}^{n} P_i(1+r/n)^{-t_i*n}$$ the price on the left side of the equation is dirty price and the price on the right hand side of the equation is clean price which is the closed price that we collected from the website. t_i represents the days from last coupon payment,n represents the number of coupon payment per year. In order to calculate the bonds' yield, we need to calculate the dirty price according to its relative clean price. Dirty price = Accrued interest + Clean Price, and Accrued interest = number of days since last coupon payment /365 * Annual Coupon rate. I calcuate the YTM of each bond for 10 days by using Excel. Then I plot the 5-year yield curve corresponding to each day of data superimposed on top of each other. The x-axis stands for the time to maturity in years, and y-axis stands for the YTM of each day. The graph explans how the yields of each day changes.



#### (b)


```{r, echo=FALSE}
library(readxl)
dirty_price_data <- read_excel("dirty price data.xlsx")
dp.matrix <- matrix(c(dirty_price_data$'2022.01.10',dirty_price_data$'2022.01.11',dirty_price_data$'2022.01.12',dirty_price_data$'2022.01.13',dirty_price_data$'2022.01.14',dirty_price_data$'2022.01.17',dirty_price_data$'2022.01.18',dirty_price_data$'2022.01.19',dirty_price_data$'2022.01.20',dirty_price_data$'2022.01.21'),nrow=10, ncol=10,byrow=TRUE)



cf_1=c(100.125)
cf_2=c(0.125,100.125)
cf_3=c(0.125,0.125,100.125)
cf_4=c(1.125,1.125,1.125,101.125)
cf_5=c(0.75,0.75,0.75,0.75,100.75)
cf_6=c(0.625,0.625,0.625,0.625,0.625,100.625)
cf_7=c(0.25,0.25,0.25,0.25,0.25,0.25,100.25)
cf_8=c(0.125,0.125,0.125,0.125,0.125,0.125,0.125,100.125)
cf_9=c(0.5,0.5,0.5,0.5,0.5,0.5,0.5,0.5,100.5)
cf_10=c(0.625,0.625,0.625,0.625,0.625,0.625,0.625,0.625,0.625,100.625)

spot_rate_matrix=matrix(nrow=10,ncol=10)


for (i in 1:10){
  srf_1 = function(x) dp.matrix[i,1]-cf_1[1]*(1+x/2)^(-2*0.5)
  s_1 = uniroot(srf_1, c(0,1))$root

  srf_2 = function(x) dp.matrix[i,2]-cf_2[1]*(1+s_1/2)^(-2*(1-0.5*1))-cf_2[2]*(1+x/2)^(-2*1)
  s_2 = uniroot(srf_2, c(0,1))$root
  
  srf_3 = function(x) dp.matrix[i,3]-cf_3[1]*(1+s_1/2)^(-2*(1.5-0.5*2))-cf_3[2]*(1+s_2/2)^(-2*(1.5-0.5*1))-cf_3[3]*(1+x/2)^(-2*1.5)
  s_3 = uniroot(srf_3, c(0,1))$root
  
  srf_4 = function(x) dp.matrix[i,4]-cf_4[1]*(1+s_1/2)^(-2*(2-0.5*3))-cf_4[2]*(1+s_2/2)^(-2*(2-0.5*2))-cf_4[3]*(1+s_3/2)^(-2*(2-0.5*1))-cf_4[4]*(1+x/2)^(-2*2)
  s_4 = uniroot(srf_4, c(0,1))$root
  
  srf_5 = function(x) dp.matrix[i,5]-cf_5[1]*(1+s_1/2)^(-2*(2.5-0.5*4))-cf_5[2]*(1+s_2/2)^(-2*(2.5-0.5*3))-cf_5[3]*(1+s_3/2)^(-2*(2.5-0.5*2))-cf_5[4]*(1+s_4/2)^(-2*(2.5-0.5*1))-cf_5[5]*(1+x/2)^(-2*2.5)
  s_5 = uniroot(srf_5, c(0,1))$root
  
  srf_6 = function(x) dp.matrix[i,6]-cf_6[1]*(1+s_1/2)^(-2*(3-0.5*5))-cf_6[2]*(1+s_2/2)^(-2*(3-0.5*4))-cf_6[3]*(1+s_3/2)^(-2*(3-0.5*3))-cf_6[4]*(1+s_4/2)^(-2*(3-0.5*2))-cf_6[5]*(1+s_5/2)^(-2*(3-0.5*1))-cf_6[6]*(1+x/2)^(-2*3)
  s_6 = uniroot(srf_6, c(0,1))$root
  
  srf_7 = function(x) dp.matrix[i,7]-cf_7[1]*(1+s_1/2)^(-2*(3.5-0.5*6))-cf_7[2]*(1+s_2/2)^(-2*(3.5-0.5*5))-cf_7[3]*(1+s_3/2)^(-2*(3.5-0.5*4))-cf_7[4]*(1+s_4/2)^(-2*(3.5-0.5*3))-cf_7[5]*(1+s_5/2)^(-2*(3.5-0.5*2))-cf_7[6]*(1+s_6/2)^(-2*(3.5-0.5*1))-cf_7[7]*(1+x/2)^(-2*3.5)
  s_7 = uniroot(srf_7, c(0,1))$root
  
 srf_8 = function(x) dp.matrix[i,8]-cf_8[1]*(1+s_1/2)^(-2*(4-0.5*7))-cf_8[2]*(1+s_2/2)^(-2*(4-0.5*6))-cf_8[3]*(1+s_3/2)^(-2*(4-0.5*5))-cf_8[4]*(1+s_4/2)^(-2*(4-0.5*4))-cf_8[5]*(1+s_5/2)^(-2*(4-0.5*3))-cf_8[6]*(1+s_6/2)^(-2*(4-0.5*2))-cf_8[7]*(1+s_7/2)^(-2*(4-0.5*1))-cf_8[8]*(1+x/2)^(-2*4)
  s_8 = uniroot(srf_8, c(0,1))$root
  
  srf_9 = function(x) dp.matrix[i,9]-cf_9[1]*(1+s_1/2)^(-2*(4.5-0.5*8))-cf_9[2]*(1+s_2/2)^(-2*(4.5-0.5*7))-cf_9[3]*(1+s_3/2)^(-2*(4.5-0.5*6))-cf_9[4]*(1+s_4/2)^(-2*(4.5-0.5*5))-cf_9[5]*(1+s_5/2)^(-2*(4.5-0.5*4))-cf_9[6]*(1+s_6/2)^(-2*(4.5-0.5*3))-cf_9[7]*(1+s_7/2)^(-2*(4.5-0.5*2))-cf_9[8]*(1+s_8/2)^(-2*(4.5-0.5*1))-cf_9[9]*(1+x/2)^(-2*4.5)
  s_9 = uniroot(srf_9, c(0,1))$root
  
  srf_10 = function(x) dp.matrix[i,10]-cf_10[1]*(1+s_1/2)^(-2*(5-0.5*9))-cf_10[2]*(1+s_2/2)^(-2*(5-0.5*8))-cf_10[3]*(1+s_3/2)^(-2*(5-0.5*7))-cf_10[4]*(1+s_4/2)^(-2*(5-0.5*6))-cf_10[5]*(1+s_5/2)^(-2*(5-0.5*5))-cf_10[6]*(1+s_6/2)^(-2*(5-0.5*4))-cf_10[7]*(1+s_7/2)^(-2*(5-0.5*3))-cf_10[8]*(1+s_8/2)^(-2*(5-0.5*2))-cf_10[9]*(1+s_9/2)^(-2*(5-0.5*1))-cf_10[10]*(1+x/2)^(-2*5)
  s_10 = uniroot(srf_10, c(0,1))$root
  
  s = rbind(s_1,s_2,s_3,s_4,s_5,s_6,s_7,s_8,s_9,s_10)
  
  for (j in 1:10){
    spot_rate_matrix[i,j]=s[j]
  }
}

year = c(0.5,1,1.5,2,2.5,3,3.5,4,4.5,5)
plot(year,spot_rate_matrix[1,],type="o", main="5-year Spot Curve", col="black", xlab="Year", ylab="Spot Rate",ylim=c(0.005,0.018))
colour = c("red","orange","yellow","green","blue","purple","pink","navy","light blue")
for (i in c(2:10)){
   lines(year,spot_rate_matrix[i,], type="o", col=colour[i-1])
}
legend("bottomright",legend=c("01/10/2022","01/11/2022","01/12/2022","01/13/2022","01/14/2022","01/17/2022","01/18/2022","01/19/2022","01/20/2022","01/21/2022"), col=c("black","red","orange","yellow","green","blue","purple","pink","navy","light blue"), lty=1, cex=0.8)

```



I use the bootstrapping technique to calculate the spot rates based on the price-yield equation using different rates for each of the 6-month terms, the equation is $$ Bond price = C_1/(1+r_1)^{1}+C_2/(1+r_2)^{2}+......+C_n/(1+r_n)^{n}+P/(1+r_n)^{n} $$ which Bond price on the left hand side is the dirty price of the bond today, C represents the cash flow, before the maturity date, bond holder receive coupon payment each 6 months in this case, on last coupon payment date, the holder can receive coupon payment plus face value of the bond. The spot rate r in the equation is annual, but the term is for half year, then we need to divide it by 2. Similarly for the power term. I list the cash flow for each bond on each coupon payment date, and use it directly in the uniroots function. Then we can get the spot rate of each bond for 10 days period. The plot with a 5-year spot curve corresponding to each day of data superimposed on-top of each other is well-labeled above.


### (c)
```{r, echo=FALSE}
spotrateyear=spot_rate_matrix[,c(2,4,6,8,10)]
fwd_matrix=matrix(nrow=10,ncol=4)

for(j in c(1:10)){
  for(i in c(1:4)){
    f1_i=function(x) (1+spotrateyear[j,i]/2)^(2*i)*(1+x/2)^(2)-(1+spotrateyear[j,i+1]/2)^(2*(i+1))
    fwd_matrix[j,i]=uniroot(f1_i,c(0,1))$root
  }
}
year=c(1,2,3,4)
plot(year,fwd_matrix[1,],type="o", main="Forward Curve", col="black", xlab="Year", ylab="Forward Rate",ylim=c(0.011,0.022))
colour = c("red","orange","yellow","green","blue","purple","pink","navy","light blue")
for (i in c(2:10)){
   lines(year,fwd_matrix[i,], type="o", col=colour[i-1])
}
legend("bottomright",legend=c("01/10/2022","01/11/2022","01/12/2022","01/13/2022","01/14/2022","01/17/2022","01/18/2022","01/19/2022","01/20/2022","01/21/2022"), col=c("black","red","orange","yellow","green","blue","purple","pink","navy","light blue"), lty=1, cex=0.8)
```

###
A forward rate is an interest rate applicable to a financial transaction that will take place in the future. We can calculate it by using spot rate on each coupon payment date. The equation is $$ 1+f_{t,1}=(1+r_{t+1})^{t+1}/(1+r_t)^{t} $$ and our coupon payment is semi-annual, then the equation will be changed to $$ (1+f_{t,1})^{2}=(1+r_{t+1}/2)^{2(t+1)}/(1+r_t/2)^{2t} $$ then I use uniroots to solve the equation. The spot rate matrix is recombined since we need 1yr-1yr,1yr-2yr,1yr-3yr,1yr-4yr, and I put four forward rates for each one of 10 bonds in one forward rate matrix. The plot with a forward curve corresponding to each day of data superimposed on-top of each other is well-labeled above.

#### 5.
```{r,echo=FALSE}
log_ytm_matrix=matrix(nrow=9,ncol=5)
for (i in c(1:5)){
  for (j in c(2:10)){
    log_ytm_matrix[j-1,i]=log(as.numeric(ytm.matrix[i,j+1])/as.numeric(ytm.matrix[i,j]))
  }
}
ytm_cov=cov(log_ytm_matrix,log_ytm_matrix)
print(ytm_cov)

log_fwd_matrix=matrix(nrow=9,ncol=4)
for (i in c(1:4)){
  for (j in c(1:9)){
    log_fwd_matrix[j,i]= log(fwd_matrix[j+1,i]/fwd_matrix[j,i])
  }
}
fwd_cov=cov(log_fwd_matrix,log_fwd_matrix)
print(fwd_cov)

```


The 5X5 co-variance matrix is for the time series of daily log-returns of yield, and 4X4 co-variance matrix is for the time series of daily log-returns of forward rates.

#### 6.

```{r, echo=FALSE}
print(eigen(ytm_cov)$values)
print(eigen(ytm_cov)$vectors)
```


The eigenvalues and eigenvectors of co-variance matrix is for the time series of daily log-returns of yield are listed above. The first eigenvalue is 2.801508e-03.


```{r, echo=FALSE}
print(eigen(fwd_cov)$values)
print(eigen(fwd_cov)$vectors)
```

The eigenvalues and eigenvectors of co-variance matrix is for the time series of daily log-returns of forward rates are listed above. The first eigenvalue is 2.385013e-03,its associated eigenvector implies the principle component since the highest eigenvalues implies the most comprehensive information included.




#### References amd GitHub Link to Code

data from: https://markets.businessinsider.com/.


GitHub Link: https://github.com/Kekekeying/APM466-A1


